---
title: "ML_Assignment_#5"
author: "Kevin Muñoz"
output: html_document
---

###SECTION 2 (GUSTO STUDY): Part 1: Logistic Regression Model

###Logistic Regression Model:


```{r, echo=TRUE}


#Clear global environment 
rm(list=ls())


#Load packages.
library(plyr)
library(dplyr)
library(pROC)
library(caret)


#Read csv file.
d.in <- read.csv("~/Desktop/gusto_data.csv")



#Change variables from continuous to categorical variables that properly represent the data. 
#Ensures all variables are properly represented as either numeric or categorical.

d.in$GROUP<- as.factor(d.in$GROUP)
d.in$SEX <- as.factor(d.in$SEX)
d.in$A65 <- as.factor(d.in$A65)
d.in$KILLIP <- as.factor(d.in$KILLIP)
d.in$SHO <- as.factor(d.in$SHO)
d.in$DIA <- as.factor(d.in$DIA)
d.in$HYP <- as.factor(d.in$HYP)
d.in$HRT <- as.factor(d.in$HRT)
d.in$ANT <- as.factor(d.in$ANT)
d.in$SMK <- as.factor(d.in$SMK)
d.in$HTN <- as.factor(d.in$HTN)
d.in$LIP <- as.factor(d.in$LIP)
d.in$PAN <- as.factor(d.in$PAN)
d.in$FAM <- as.factor(d.in$FAM)
d.in$ST4 <- as.factor(d.in$ST4)
d.in$TTR <- as.factor(d.in$TTR)
d.in$HIG <- as.factor(d.in$HIG)
d.in$PMI <- as.factor(d.in$PMI)
d.in$DAY30 <- as.factor(d.in$DAY30)


#Split the dataset into training (groups = sample2 or sample4 or sample5) and testing (group = west). 


train_sample_groups_2_4_5 <- filter(d.in, GROUP != "west")

summary(train_sample_groups_2_4_5)

test_west_group <- filter(d.in, GROUP == "west")

summary(test_west_group)


###Logistic Regression Framework with all features to predict 30-day mortality in a logistic regression framework.


#Create labels and create a logistic regression training model  
levels(train_sample_groups_2_4_5$DAY30) <- c("No", "Yes")
fitControl <- trainControl(method = "cv", number = 5)
glm.model <- train(DAY30 ~ . ,
                   data = train_sample_groups_2_4_5,
                   method = "glm",
                   family = "binomial",
                   trControl = fitControl)


print(glm.model)


#Apply to test dataset by creating a prediction object.
preds <- predict(object = glm.model, test_west_group, type = "prob")
test_west_group$predicted_prob_DAY30 <- preds$Yes
glm_perf <- roc(response = test_west_group$DAY30,
                predictor = test_west_group$predicted_prob_DAY30)



#Get performance: calculate AUC and their confidence intervals.

print(pROC::auc(glm_perf))
print(pROC::ci.auc(glm_perf))

```


###SECTION 2 (GUSTO STUDY): Part 2: Ridge Regression

###Ridge Regression:

```{r, echo=TRUE}




#Clear environment.
rm(list=ls())


#Load packages.
library(plyr)
library(dplyr)
library(pROC)
library(caret)


#Read csv file.
d.in <- read.csv("~/Desktop/gusto_data.csv")



#Change variables from continuous to categorical variables that properly represent the data. 
#This will help with question 3 to ensure all variables are properly represented as either numeric or categorical.

d.in$GROUP<- as.factor(d.in$GROUP)
d.in$SEX <- as.factor(d.in$SEX)
d.in$A65 <- as.factor(d.in$A65)
d.in$KILLIP <- as.factor(d.in$KILLIP)
d.in$SHO <- as.factor(d.in$SHO)
d.in$DIA <- as.factor(d.in$DIA)
d.in$HYP <- as.factor(d.in$HYP)
d.in$HRT <- as.factor(d.in$HRT)
d.in$ANT <- as.factor(d.in$ANT)
d.in$SMK <- as.factor(d.in$SMK)
d.in$HTN <- as.factor(d.in$HTN)
d.in$LIP <- as.factor(d.in$LIP)
d.in$PAN <- as.factor(d.in$PAN)
d.in$FAM <- as.factor(d.in$FAM)
d.in$ST4 <- as.factor(d.in$ST4)
d.in$TTR <- as.factor(d.in$TTR)
d.in$HIG <- as.factor(d.in$HIG)
d.in$PMI <- as.factor(d.in$PMI)
d.in$DAY30 <- as.factor(d.in$DAY30)


#Split the dataset into training (groups = sample2 or sample4 or sample5) and testing (group = west). 


train_sample_groups_2_4_5 <- filter(d.in, GROUP != "west")

summary(train_sample_groups_2_4_5)

test_west_group <- filter(d.in, GROUP == "west")

summary(test_west_group)


#Logistic Regression Framework with all features to predict 30-day mortality. Also, with regularization on the L2 term. 
#Utilization of 5 fold cross validation to build the parameters for ridge regression model.


#Load package.
library(glmnetUtils)

#Create labels and ridge regression model. 
levels(train_sample_groups_2_4_5$DAY30) <- c("No", "Yes")

m_ridge <- glmnet(DAY30 ~ . ,
                  data = train_sample_groups_2_4_5,
                  family = "binomial",
                  alpha = 0)

#Weights converge by plotting.
plot(m_ridge, xvar = "lambda")


#For a good range lambda, we need to use CV on glmnet first then fine tune it later with the package caret. 


m_ridge_cv <- cv.glmnet(DAY30 ~ . ,
                        data = train_sample_groups_2_4_5,
                        family = "binomial",
                        alpha = 0,
                        nfolds = 5)

plot(m_ridge_cv)


#Minimum lambda.

m_ridge_cv$lambda.min


#Minimum lambda plus one SD.

m_ridge_cv$lambda.1se

#Create a grid for finding the perfect lambda and alpha. Also, everything between.

min_lambda <- min(m_ridge_cv$lambda.min, m_ridge_cv$lambda.1se)
max_lambda <- max(m_ridge_cv$lambda.min, m_ridge_cv$lambda.1se)
lambda_grid <- seq(min_lambda, max_lambda, by = 0.001)


alpha_grid <- c(0)
objGrid <- expand.grid(alpha = alpha_grid,
                       lambda = lambda_grid)


#Regularization in R: 
#Create labels and a logistic regression training model using cross-validation.  
levels(train_sample_groups_2_4_5$DAY30) <- c("No", "Yes")
fitControl <- trainControl(method = "cv", number = 5)
reg.model <- train(DAY30 ~ . ,
                   data = train_sample_groups_2_4_5,
                   method = "glmnet",
                   tuneGrid = objGrid,
                   trControl = fitControl)



print(reg.model)


#Apply to test dataset by creating a prediction object. 
preds <- predict(object = reg.model, test_west_group, type = "prob")
test_west_group$predicted_prob_DAY30 <- preds$Yes
reg_perf <- roc(response = test_west_group$DAY30,
                predictor = test_west_group$predicted_prob_DAY30)


#Get performance: calculate AUC and their confidence intervals.

print(pROC::auc(reg_perf))
print(pROC::ci.auc(reg_perf))

```




###SECTION 2 (GUSTO STUDY): Part 3: ANN Model

###ANN Model:


```{r, echo=TRUE}



#Clear environment.
rm(list=ls())


#Load packages.
library(plyr)
library(dplyr)
library(pROC)
library(caret)


#Read csv file.
d.in <- read.csv("~/Desktop/gusto_data.csv")



#Change variables from continuous to categorical variables that properly represent the data. 
#This will help with question 3 to ensure all variables are properly represented as either numeric or categorical.

d.in$GROUP<- as.factor(d.in$GROUP)
d.in$SEX <- as.factor(d.in$SEX)
d.in$A65 <- as.factor(d.in$A65)
d.in$KILLIP <- as.factor(d.in$KILLIP)
d.in$SHO <- as.factor(d.in$SHO)
d.in$DIA <- as.factor(d.in$DIA)
d.in$HYP <- as.factor(d.in$HYP)
d.in$HRT <- as.factor(d.in$HRT)
d.in$ANT <- as.factor(d.in$ANT)
d.in$SMK <- as.factor(d.in$SMK)
d.in$HTN <- as.factor(d.in$HTN)
d.in$LIP <- as.factor(d.in$LIP)
d.in$PAN <- as.factor(d.in$PAN)
d.in$FAM <- as.factor(d.in$FAM)
d.in$ST4 <- as.factor(d.in$ST4)
d.in$TTR <- as.factor(d.in$TTR)
d.in$HIG <- as.factor(d.in$HIG)
d.in$PMI <- as.factor(d.in$PMI)
d.in$DAY30 <- as.factor(d.in$DAY30)


#Split the dataset into training (groups = sample2 or sample4 or sample5) and testing (group = west). 


train_sample_groups_2_4_5 <- filter(d.in, GROUP != "west")

summary(train_sample_groups_2_4_5)

test_west_group <- filter(d.in, GROUP == "west")

summary(test_west_group)



#Balancing classes using SMOTE in training dataset. 

#train_sample_groups_2_4_5 <- SMOTE(DAY30 ~ ., data = train_sample_groups_2_4_5,
                 #perc.over = 200, perc.under = 100, k = 1)



#Normalization and standardization of continuous variables. 
#Use training dataset and subtract mean to center, divide by standard deviation
preProcValues <- preProcess(train_sample_groups_2_4_5, 
                            method = c("center", "scale"))


trainTransformed <- predict(preProcValues, train_sample_groups_2_4_5)
testTransformed <- predict(preProcValues, test_west_group)


#set seed for reproducibility. 
set.seed(456)

#Create labels and create a ANN model.
levels(trainTransformed$DAY30) <- c("No", "Yes")

#Choosing a grid size for the hidden layer and the decay. 

objGrid <- expand.grid(size = seq(2, 50, by =1), decay = c(0.001))

fitControl <- trainControl(method = "cv", number = 5, classProbs = TRUE)
ann.model <- train(DAY30~ . ,
                   data = trainTransformed,
                   method = "nnet",
                   metric = "AUC",
                   trControl = fitControl,
                   tuneGrid = objGrid,
                   trace = FALSE)

print(ann.model)


#Apply to test dataset by creating a prediction object.
preds <- predict(object = ann.model, testTransformed, type = "prob")
testTransformed$predicted_prob_DAY30 <- preds$Yes
ann_perf <- pROC::roc(response = testTransformed$DAY30,
                      predictor = testTransformed$predicted_prob_DAY30)


#Get performance: calculate AUC and their confidence intervals.

print(pROC::auc(ann_perf))
print(pROC::ci.auc(ann_perf))


```


###SECTION 2 (GUSTO STUDY): Part 4: Random Forest Model

### Random Forest Model:


```{r, echo=TRUE}

#Clear environment.
rm(list=ls())


#Load packages.
library(plyr)
library(dplyr)
library(pROC)
library(caret)


#Read csv file.
d.in <- read.csv("~/Desktop/gusto_data.csv")



#Change variables from continuous to categorical variables that properly represent the data. 
#This will help with question 3 to ensure all variables are properly represented as either numeric or categorical.

d.in$GROUP<- as.factor(d.in$GROUP)
d.in$SEX <- as.factor(d.in$SEX)
d.in$A65 <- as.factor(d.in$A65)
d.in$KILLIP <- as.factor(d.in$KILLIP)
d.in$SHO <- as.factor(d.in$SHO)
d.in$DIA <- as.factor(d.in$DIA)
d.in$HYP <- as.factor(d.in$HYP)
d.in$HRT <- as.factor(d.in$HRT)
d.in$ANT <- as.factor(d.in$ANT)
d.in$SMK <- as.factor(d.in$SMK)
d.in$HTN <- as.factor(d.in$HTN)
d.in$LIP <- as.factor(d.in$LIP)
d.in$PAN <- as.factor(d.in$PAN)
d.in$FAM <- as.factor(d.in$FAM)
d.in$ST4 <- as.factor(d.in$ST4)
d.in$TTR <- as.factor(d.in$TTR)
d.in$HIG <- as.factor(d.in$HIG)
d.in$PMI <- as.factor(d.in$PMI)
d.in$DAY30 <- as.factor(d.in$DAY30)


#Split the dataset into training (groups = sample2 or sample4 or sample5) and testing (group = west). 


train_sample_groups_2_4_5 <- filter(d.in, GROUP != "west")

summary(train_sample_groups_2_4_5)

test_west_group <- filter(d.in, GROUP == "west")

summary(test_west_group)



#Set parameters:
#Take the square root of number of features and create grid by setting grid to be +/- 2.  

p_mtry <- round(sqrt(ncol(train_sample_groups_2_4_5)))
mtry_grid <- seq(p_mtry -2, p_mtry +2, by = 1)
objGrid <- expand.grid(.mtry = mtry_grid)



#Create labels, cross 5-fold validation, and create a random forest training model.
#Can change the number of trees to compare the performance output of all the models like 500 or 1000. 

levels(train_sample_groups_2_4_5$DAY30) <- c("No", "Yes")
fitControl <- trainControl(method = "cv", number = 5)



rf.model <- train(DAY30 ~ . ,
                  data = train_sample_groups_2_4_5,
                  method = "rf",
                  tuneGrid = objGrid,
                  ntree = 1000, 
                  trControl = fitControl,
                  importance= TRUE)


library(ggplot2)

#Use varimp function from caret package to tell across all variables for a single model which are most important. 
feat_imp <- caret::varImp(rf.model, scale=F)

feat_imp

#Use ggplot to plot variable importance graph. 
p <- ggplot(feat_imp) + ggtitle("Variable Importance Plot")

p <- p + labs(y = "Importance", x = "Variable Features") 

p


#Apply to test dataset by creating a prediction object.
preds <- predict(object = rf.model, test_west_group, type = "prob")
test_west_group$predicted_prob_DAY30 <- preds$Yes
rf_perf <- roc(response = test_west_group$DAY30,
               predictor = test_west_group$predicted_prob_DAY30)



#Get performance: calculate AUC and their confidence intervals.

print(pROC::auc(rf_perf))
print(pROC::ci.auc(rf_perf))



```




###SECTION 2 (GUSTO STUDY): Part 5: Gradient Boosting Model

### Gradient Boosting Model:



```{r, echo=TRUE}

#Clear environment.
rm(list=ls())


#Load packages.
library(plyr)
library(dplyr)
library(pROC)
library(caret)


#Read csv file.
d.in <- read.csv("~/Desktop/gusto_data.csv")



#Change variables from continuous to categorical variables that properly represent the data. 
#This will help with question 3 to ensure all variables are properly represented as either numeric or categorical.

d.in$GROUP<- as.factor(d.in$GROUP)
d.in$SEX <- as.factor(d.in$SEX)
d.in$A65 <- as.factor(d.in$A65)
d.in$KILLIP <- as.factor(d.in$KILLIP)
d.in$SHO <- as.factor(d.in$SHO)
d.in$DIA <- as.factor(d.in$DIA)
d.in$HYP <- as.factor(d.in$HYP)
d.in$HRT <- as.factor(d.in$HRT)
d.in$ANT <- as.factor(d.in$ANT)
d.in$SMK <- as.factor(d.in$SMK)
d.in$HTN <- as.factor(d.in$HTN)
d.in$LIP <- as.factor(d.in$LIP)
d.in$PAN <- as.factor(d.in$PAN)
d.in$FAM <- as.factor(d.in$FAM)
d.in$ST4 <- as.factor(d.in$ST4)
d.in$TTR <- as.factor(d.in$TTR)
d.in$HIG <- as.factor(d.in$HIG)
d.in$PMI <- as.factor(d.in$PMI)
d.in$DAY30 <- as.factor(d.in$DAY30)


#Split the dataset into training (groups = sample2 or sample4 or sample5) and testing (group = west). 


train_sample_groups_2_4_5 <- filter(d.in, GROUP != "west")

summary(train_sample_groups_2_4_5)

test_west_group <- filter(d.in, GROUP == "west")

summary(test_west_group)


#set seed for reproducibility. 
set.seed(456)
p_n.trees <- c(500, 1000) #Number of trees.
p_lr <- c(0.01, 0.05) #Parameter that needs to be tuned is the learning rate or shrinkage.
p_interaction_depth <- c(2, 3, 5) #Max number of splits per tree (want to create a small tree). 
p_min_obs <- c(10) #Minimum observations for tree to stop growing.


#Created ObjGrid with four variables and different observations.
objGrid <- expand.grid(interaction.depth = p_interaction_depth,
                       n.trees = p_n.trees,
                       shrinkage = p_lr,
                       n.minobsinnode = p_min_obs)



levels(train_sample_groups_2_4_5$DAY30) <- c("No", "Yes")
fitControl <- trainControl(method = "cv", number = 5)


#Create labels, cross 5-fold validation, and create a gbm training model.
#Another way to the output of gbm because gbm doesn't suppress the output.
temp <- capture.output(gbm.model <- train(DAY30 ~ . ,
                                          data = train_sample_groups_2_4_5,
                                          method = "gbm",
                                          tuneGrid = objGrid,
                                          trControl = fitControl))


print(gbm.model)



#Apply to test dataset by creating a prediction object.
preds <- predict(object = gbm.model, test_west_group, type = "prob")
test_west_group$predicted_prob_DAY30 <- preds$Yes
gbm_perf <- roc(response = test_west_group$DAY30,
               predictor = test_west_group$predicted_prob_DAY30)



#Get performance: calculate AUC and their confidence intervals.

print(pROC::auc(gbm_perf))
print(pROC::ci.auc(gbm_perf))



```




###SECTION 2 (GUSTO STUDY): Part 6: SVM Model

### Support Vector Model:


```{r, echo=TRUE}




#Clear environment.
rm(list=ls())


#Load packages.
library(plyr)
library(dplyr)
library(pROC)
library(caret)


#Read csv file.
d.in <- read.csv("~/Desktop/gusto_data.csv")



#Change variables from continuous to categorical variables that properly represent the data. 
#This will help with question 3 to ensure all variables are properly represented as either numeric or categorical.

d.in$GROUP<- as.factor(d.in$GROUP)
d.in$SEX <- as.factor(d.in$SEX)
d.in$A65 <- as.factor(d.in$A65)
d.in$KILLIP <- as.factor(d.in$KILLIP)
d.in$SHO <- as.factor(d.in$SHO)
d.in$DIA <- as.factor(d.in$DIA)
d.in$HYP <- as.factor(d.in$HYP)
d.in$HRT <- as.factor(d.in$HRT)
d.in$ANT <- as.factor(d.in$ANT)
d.in$SMK <- as.factor(d.in$SMK)
d.in$HTN <- as.factor(d.in$HTN)
d.in$LIP <- as.factor(d.in$LIP)
d.in$PAN <- as.factor(d.in$PAN)
d.in$FAM <- as.factor(d.in$FAM)
d.in$ST4 <- as.factor(d.in$ST4)
d.in$TTR <- as.factor(d.in$TTR)
d.in$HIG <- as.factor(d.in$HIG)
d.in$PMI <- as.factor(d.in$PMI)
d.in$DAY30 <- as.factor(d.in$DAY30)


#Split the dataset into training (groups = sample2 or sample4 or sample5) and testing (group = west). 


train_sample_groups_2_4_5 <- filter(d.in, GROUP != "west")

summary(train_sample_groups_2_4_5)

test_west_group <- filter(d.in, GROUP == "west")

summary(test_west_group)



set.seed(892)

#Setup for cross validation.

#Tune for the cost parameter c. 
objGrid <- expand.grid(C = c(0.01, 0.1, 1, 10, 100, 1000))

levels(train_sample_groups_2_4_5$DAY30) <- c("No", "Yes")
fitControl <- trainControl(method = "cv", number = 5, classProbs = TRUE)

m1.svm <- train(DAY30 ~ . ,
                  data = train_sample_groups_2_4_5,
                  method = "svmLinear",
                  tuneGrid = objGrid,
                  trControl = fitControl)

m1.svm


#Apply to test dataset by creating a prediction object.
preds <- predict(object = m1.svm, test_west_group, type = "prob")
test_west_group$predicted_prob_DAY30 <- preds$Yes
svm_perf <- roc(response = test_west_group$DAY30,
                predictor = test_west_group$predicted_prob_DAY30)



#Get performance: calculate AUC and their confidence intervals.

print(pROC::auc(svm_perf))
print(pROC::ci.auc(svm_perf))




```









